Relationship between solar wind, $D_{st}$, and plasmasphere mass density on on-hour time scales

### 2015-06-02 paper.tex

\twocolumn[
  \begin{@twocolumnfalse}
\maketitle
\hrule
\begin{abstract}
We consider solar wind conditions surrounding time intervals of large negative excursions of the $D_{st}$ index and large positive excursions of plasmaspheric mass density. Consistent with previous results, a sudden decrease in $D_{st}$ is associated with an increase in equatorial mass density.  We find ...
\vspace{2em}
\end{abstract}
 \end{@twocolumnfalse}
  ]

@@

%This paper intends to investigate solar wind and magnetospheric conditions immediately before geomagnetic storm onset to determine trends and useful correlations, specifically looking at the effects on equatorial mass density ($\rho_{eq}$).
In this paper we test the relationship of short timescale effects on mass density from $D_{ST}$, $F_{10.7}$, and $B_Z$, in comparison to the long time scales presented in \cite{Takahashi2010,Takahashi2006}, as well as looking at other preconditioning factors by analysis of hourly data within a day surrounding storm onsets.

%Keep the following for now.  Not sure where we will put it.
%We also hope that by looking at all possible storms under less stringent qualifications, instead of hand selected storms with specific selection criteria, our results will be more generalizable and less prone to bias.

\cite{Takahashi2006} considered the relationship between geomagnetic activity and average ion mass ($M$).  A scatter plot of log$_{10}(M)$ versus $-D_{st}$ showed an increasing trend.  The scatter plot of log$_{10}(M)$ versus $K_p$ had near zero correlation, but a positive trend was found when log$_{10}(M)$ versus 1.5-day averaged $K_p$ was considered, and this trend persisted when 3-day averages were used.

\cite{Takahashi2010} found that spikes in the Disturbance Storm Time ($D_{ST}$) index coincide with with significant changes in $\rho_{eq}$ at an L-shell of 6.8$R_E$. For five storms over a 20 day period two had $\rho_{eq}$ spikes after the $D_{ST}$ drop, two had $\rho_{eq}$ spikes before the drop, and one showed little change in $\rho_{eq}$. They then show an epoch analysis where $\rho_{eq}$ is seen to spike the day of a $D_{ST}$ drop, using a daily average of 30 minute $\rho_{eq}$ and one hour $D_{ST}$ measurements.

>>>>

\twocolumn[
  \begin{@twocolumnfalse}
\maketitle
\hrule
\begin{abstract}
This paper looks at how various magnetosphere conditions behave around the onset of geomagnetic storms. It confirms results from previous papers that a sudden drop in $D_{ST}$ correlates with a spike in equatorial mass density, while adding a level of depth and specification to the applicability of their results. It compares data at 1-hour averages to data at 1-day averages, and briefly to data at 27-day averages to examine if the trend holds at varying time scales and finds that it does, so long as the data is pre-processed to linearly remove a large time scale $F_{10.7}$ dependence. 
\vspace{2em}
\end{abstract}
 \end{@twocolumnfalse}
  ]

@@

\cite{Takahashi2010} state that spikes in the Disturbance Storm Time ($D_{ST}$) index coincide with with significant changes in $\rho_{eq}$ at an L-shell of 6.8$R_E$. For five storms over a 20 day period two had $\rho_{eq}$ spikes after the $D_{ST}$ drop, two had $\rho_{eq}$ spikes before the drop, and one showed little change in $\rho_{eq}$. They then show an epoch analysis where $\rho_{eq}$ is seen to spike the day of a $D_{ST}$ drop, using a daily average of 30 minute $\rho_{eq}$ and one hour $D_{ST}$ measurements.

The prior paper, \cite{Takahashi2006}, show how trends in storm dependence on mass/density only appear in longer timescales. Looking at $Kp$ vs mass ($amu$) in the range of 6 to 7 $R_E$, a trend shows up in the 1.5 day averages that doesn't appear in the 3 hour averages.
